Total Score

### 2023-04-20 src/MSAtotal_scores.Rmd
github check
cleaning up code

---
title: "Total Score"
author: "Grace Bianchi"
date: '2022-11-14'
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

@@

```{r}
# read in the data
real_estate_score_data <- read_sheet("https://docs.google.com/spreadsheets/d/1Ssmb7pUd4NFZocVp9C75SRMOV6tAJCCt0TjrCYBB8pY/edit#gid=1038418521") %>% 
  select(cbsa, real_estate_score)

@@

```{r read in data}
criteria_all <- real_estate_score_data %>% 
  inner_join(rental_policy_score_data)  %>% 
  inner_join(reopt_data) %>% 
  inner_join(climate_risk_data) # combine all criteria scores into one dataframe

@@

# rename columns
criteria_all <- criteria_all[, col_order] %>% 
  rename(financial_score = normalized_irr,
         co2_score = inverted_n_total_co2,
         health_impact_score = inverted_n_health_emissions_cost,
         electricity_score = normalized_annual_renewable,
         climate_risk_score = normalized_inverted_climate_risk) %>% 
  mutate(real_estate_score = as.numeric(real_estate_score),
         landlord_score = as.numeric(landlord_score))

@@

```{r FINAL student weights}
# weight criteria for students
criteria_weights_students <- c(0.18, 0.0, 0.10, 0.19, 0.17, 0.14, 0.22)

#make landlord_score column
criteria_all <- criteria_all %>% 
  mutate(equity_scenario_score = cbsa,
         client_FINAL_score = cbsa) #this is just a place holder

for (i in 1:nrow(criteria_all)) {
  cbsa_data <- c(criteria_all$real_estate_score[i],
                criteria_all$landlord_score[i],
                 criteria_all$financial_score[i],
                 criteria_all$co2_score[i],
                criteria_all$health_impact_score[i],
                criteria_all$electricity_score[i],
                criteria_all$climate_risk_score[i]
                  )
  criteria_all$equity_scenario_score[i] = weighted.mean(cbsa_data, criteria_weights_students)
}

```{r Client FINAL WEIGHTS}

client_weights <- c(0.18, 0.07, 0.11, 0.07, 0.0, 0.07, 0.50)

# multiply column by respective weight and sum for total

for (i in 1:nrow(criteria_all)) {
  cbsa_data <- c(criteria_all$real_estate_score[i],
                criteria_all$landlord_score[i],
                 criteria_all$financial_score[i],
                 criteria_all$co2_score[i],
                criteria_all$health_impact_score[i],
                criteria_all$electricity_score[i],
                criteria_all$climate_risk_score[i]
                  )
  criteria_all$client_FINAL_score[i] = weighted.mean(cbsa_data, client_weights) 
}

```{r}

criteria_all <- criteria_all %>% 
  mutate_if(is.numeric,
            round,
            digits = 5)

# creating a function with criteria data and weight inputs
calculate_wt_criteria <- function(data, weights){
    wt_criteria <- data %>%
      # multiply each criteria by respective weight
      mutate(wt_real_estate = real_estate_score * weights[1],
             wt_landlord_policy = landlord_score * weights[2], 
             wt_solar_irr = financial_score * weights[3],
            wt_co2 = co2_score * weights[4], 
             wt_health_impacts = health_impact_score * weights[5],
             wt_electricity = electricity_score * weights[6],
             wt_climate_risk = climate_risk_score * weights[7]) %>% 
      # select weighted columns and rename them
      select(!c(ends_with("_score"))) %>%
      # create column for investment score
      mutate(investment_score = 0) %>%
      # for loop to sum weighted columns for each row
      mutate(investment_score = apply(.[, 4:10], 1, sum)) %>% 
      rename("Real Estate"="wt_real_estate",
             "Landlord Policy"="wt_landlord_policy",
             "Electricity Production"="wt_electricity", 
             "CO2 Emission Reduction"="wt_co2",
             "Climate Risk Avoided"="wt_climate_risk",
             "Health Impacts"="wt_health_impacts", 
             "Solar IRR"="wt_solar_irr") 
 }

@@

# equity weights
wt_criteria_equity <- calculate_wt_criteria(criteria_all, criteria_weights_students) %>% 
  unite("msa_state", city_msa:state, sep = ", ")

>>>>

---
title: "Total Score"
author: "Grace Bianchi"
date: '2022-11-14'
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=TRUE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

@@

```{r}
# read in the normalized data
real_estate_score_data <- read_sheet("https://docs.google.com/spreadsheets/d/1Ssmb7pUd4NFZocVp9C75SRMOV6tAJCCt0TjrCYBB8pY/edit#gid=1038418521") %>% 
  select(cbsa, real_estate_score)

@@

```{r data wrangling}
criteria_all <- real_estate_score_data %>% 
  inner_join(rental_policy_score_data)  %>% 
  inner_join(reopt_data) %>% 
  inner_join(climate_risk_data) # combine all criteria scores into one dataframe

@@

# rename columns
criteria_all <- criteria_all[, col_order] %>% 
  rename(financial_score = normalized_irr,
         co2_score = inverted_n_total_co2,
         health_impact_score = inverted_n_health_emissions_cost,
         electricity_score = normalized_annual_renewable,
         climate_risk_score = normalized_inverted_climate_risk) %>% 
  mutate(real_estate_score = as.numeric(real_estate_score),
         landlord_score = as.numeric(landlord_score)) %>%
  # round criteria scores
  mutate_if(is.numeric,round, digits = 5)

@@

```{r model weights}
# weights for each criteria must be in this order
equity_weights <- c(0.18, # real estate weight
                    0.0, # landlord weight
                    0.10, # financial weight
                    0.19, # co2 emissions avoided weight
                    0.17, # health impacts weight
                    0.14, # electricity weight
                    0.22) # climate risk weight

client_weights <- c(0.18, 0.07, 0.11, 0.07, 0.0, 0.07, 0.50)
```

# error check so weights sum to 1
# incorporate shiny??

```{r}
# creating a function with criteria data and weight inputs
calculate_wt_criteria <- function(data, weights){
    wt_criteria <- data %>%
      # multiply each criteria by respective weight
      mutate(wt_real_estate = real_estate_score * weights[1],
             wt_landlord_policy = landlord_score * weights[2], 
             wt_solar_irr = financial_score * weights[3],
            wt_co2 = co2_score * weights[4], 
             wt_health_impacts = health_impact_score * weights[5],
             wt_electricity = electricity_score * weights[6],
             wt_climate_risk = climate_risk_score * weights[7]) %>% 
      # select weighted columns and rename them
      select(!c(ends_with("_score"))) %>%
      # create column for investment score
      mutate(investment_score = 0) %>%
      # for loop to sum weighted columns for each row
      mutate(investment_score = apply(.[, 4:10], 1, sum)) %>% 
      rename("Real Estate"="wt_real_estate",
             "Landlord Policy"="wt_landlord_policy",
             "Electricity Production"="wt_electricity", 
             "CO2 Emission Reduction"="wt_co2",
             "Climate Risk Avoided"="wt_climate_risk",
             "Health Impacts"="wt_health_impacts", 
             "Solar IRR"="wt_solar_irr") %>%  
      unite("msa_state", city_msa:state, sep = ", ")
 }

@@

# equity weights
wt_criteria_equity <- calculate_wt_criteria(criteria_all, equity_weights) %>% 
  unite("msa_state", city_msa:state, sep = ", ")
